### 2016-01-25 inputs/materials_and_methods.tex
mainly supplementary, plus small changes all over

The 715 samples from alfalfa (Alfalfa-Med and Alfalfa-PV) and rice (Rice) populations were genotyped using the genotyping-by-sequencing (GBS) approach \cite{elshire_robust_2011}. Different GBS protocols were used to genotype the alfalfa and rice samples, since genotyping was carried out under different projects.\\
In alfalfa, DNA was isolated from fresh leaf tissues by the Wizard~\textregistered~Genomic DNA Purification Kit (Promega, A1125) and quantified with a Quant-iT PicoGreen dsDNA assay kit (Life Technologies, P7589). Two libraries were constructed for the two populations, where 100 ng of each DNA was digested with ApeKI (NEB, R0643L) and then ligated to a unique barcoded adapter and a common adapter. For the reference population Alfalfa-Med, 5 nmoles each of the primers and NEB 2X Taq Master Mix (NEB Cat \# M0270S) were included in the PCR reaction according to \cite{elshire_robust_2011} original protocol, while for Alfalfa-PV the KAPA library amplification readymix (Kapa Biosystems Cat \# KK2611) was used instead. Each library was sequenced in two lanes on Illumina HiSeq 2000 at the Genomic Sequencing and Analysis Facility at the University of Texas at Austin, Texas, USA. For both populations post sequencing analysis and SNP calling was carried out using Tassel UNEAK pipeline \cite{lu_switchgrass_2013}. A further quality filter, implemented through ad-hoc Python scripts, removed heterozygous loci with less than 4 and homozygous loci with less than 11 aligned reads. Alfalfa is an autotetraploid plant species and can therefore present three different heterozygous genotypes: AAAB, AABB and ABBB. Following \cite{li_saturated_2014} and \cite{li_genomic_2015}, these were all considered diploid heterozygotes (i.e. AB), while the two homozygous genotypes (AAAA and BBBB) were considered diploid homozygotes (i.e., AA or BB).\\
In rice, DNA was isolated from three-weeks old leaves using the DNeasy Plant Mini Kit (QIAGEN) with a TECAN Freedom EVO150 liquid handling robot (TECAN Group Ltd, Switzerland). DNA digestion was performed on 100 ng samples in 96-well plates using ApeKI, which was shown to cut every 1 kb on average in a in-silico digestion of the Nipponbare reference genome, and 96-plex libraries constructed according to the GBS protocol. The libraries were loaded into Genome Analyzer II (Illumina, Inc., San Diego, CA) for sequencing. Raw sequence data filtering, sequence alignment to the rice reference genome (Os-Nipponbare-Reference-IRGSP-1.0, \cite{kawahara_improvement_2013}), and SNP calling from GBS genotyping were carried out using the Tassel GBS pipeline \cite{glaubitz_tassel-gbs:_2014}.\\
In total, 32\,706, 40\,734 and 166\,418 SNP markers were called by GBS in Alfalfa-PV, Alfalfa-Med and Rice, respectively (see Table~\ref{tab:descriptive_statistics}). In both datasets SNP variants were renamed so that AA represented the most common homozygote, and BB the least common homozygote.

>>>>

The 715 samples from alfalfa (Alfalfa-Med and Alfalfa-PV) and rice (Rice) populations were genotyped using the genotyping-by-sequencing (GBS) approach \cite{elshire_robust_2011}. Different GBS protocols were used to genotype the alfalfa and rice samples, since genotyping was carried out under different projects.\\
In alfalfa, DNA was isolated from fresh leaf tissues by the Wizard~\textregistered~Genomic DNA Purification Kit (Promega, A1125) and quantified with a Quant-iT PicoGreen dsDNA assay kit (Life Technologies, P7589). Two libraries were constructed for the two populations, where 100 ng of each DNA was digested with ApeKI (NEB, R0643L) and then ligated to a unique barcoded adapter and a common adapter. For the reference population Alfalfa-Med, 5 nmoles each of the primers and NEB 2X Taq Master Mix (NEB Cat \# M0270S) were included in the PCR reaction according to \cite{elshire_robust_2011} original protocol, while for Alfalfa-PV the KAPA library amplification readymix (Kapa Biosystems Cat \# KK2611) was used instead. Each library was sequenced in two lanes on Illumina HiSeq 2000 at the Genomic Sequencing and Analysis Facility at the University of Texas at Austin, Texas, USA. For both populations post sequencing analysis and SNP calling was carried out using Tassel UNEAK pipeline \cite{lu_switchgrass_2013}.

A further quality filter, implemented through ad-hoc Python scripts, removed heterozygous loci with less than 4 and homozygous loci with less than 11 aligned reads. A similar filtering was performed in \cite{Rocher_validation_2015} using less restrictive thresholds.

. Alfalfa is an autotetraploid plant species and can therefore present three different heterozygous genotypes: AAAB, AABB and ABBB. Following \cite{li_saturated_2014} and \cite{li_genomic_2015}, these were all considered diploid heterozygotes (i.e. AB), while the two homozygous genotypes (AAAA and BBBB) were considered diploid homozygotes (i.e., AA or BB).\\

In rice, DNA was isolated from three-weeks old leaves using the DNeasy Plant Mini Kit (QIAGEN) with a TECAN Freedom EVO150 liquid handling robot (TECAN Group Ltd, Switzerland). DNA digestion was performed on 100 ng samples in 96-well plates using ApeKI, which was shown to cut every 1 kb on average in a in-silico digestion of the Nipponbare reference genome, and 96-plex libraries constructed according to the GBS protocol. The libraries were loaded into Genome Analyzer II (Illumina, Inc., San Diego, CA) for sequencing. Raw sequence data filtering, sequence alignment to the rice reference genome (Os-Nipponbare-Reference-IRGSP-1.0, \cite{kawahara_improvement_2013}), and SNP calling from GBS genotyping were carried out using the Tassel GBS pipeline \cite{glaubitz_tassel-gbs:_2014}.\\
In total, 32\,706, 40\,734 and 166\,418 SNP markers were called by GBS in Alfalfa-PV, Alfalfa-Med and Rice, respectively (see Table~\ref{tab:descriptive_statistics}). In both datasets SNP variants were renamed so that AA represented the most common homozygote, and BB the least common homozygote.

### 2016-01-25 inputs/supplementary.tex
other references
mainly supplementary, plus small changes all over

>>>>

\subsection{SNP calling thresholds}
\label{sec:SNP_calling_thresholds}
Genotype calling can incur in two types of error, namely a heterozygote wrongly identified as homozigote ($HET \rightarrow HOM$) and the reverse case ($HOM \rightarrow HET$). 
Under the simplifying hypothesis of absence of sequencing errors and homogeneous sequencing quality, it is possible to estimate error rates for a tetraploid genome using the following formula:

\begin{equation}
\begin{cases}
0.75^{N_h} + 0.25^{N_h} & HET_{3-1} \rightarrow HOM \\ 
2 \times 0.5^{N_h}      & HET_{2-2} \rightarrow HOM \\ 
0                       & HOM \rightarrow HET
\end{cases}
\end{equation}

Where $N_h$ is the minimum number of reads required to label a stack of aligning reads as heterozygote. The discussed case of $N_h = 11$ brings to a total expected $HET \rightarrow HOM$ error rate of 4.22\%.\\
The hypothesis mentioned above greatly simplyfies the real case (e.g. $HOM \rightarrow HET$ errors can not happen, since a homozigote will always produce 100\% of reads of the same type). When accounting sequencing errors and reads quality a term for $HOM \rightarrow HET$ appears, due to reads sequences as the wrong allele. In the presented study we thus applied a threshold of 4 required reads per stack to call a homozigote to minimize this kind of error.

\subsection{Supplementary figures}
\label{sec:supplementary_figures}
